### 2025-01-04 dk_example/example.tex
start adding the dk example

\documentclass[a4paper,11pt]{article}

% load packages
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array}
\usepackage{mathtools}
\usepackage{bbm}

% define own colors and use colored text
\usepackage{xcolor}

\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{positioning}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}
\usetikzlibrary{backgrounds}

%%% standard additional commands
%epsilon
\newcommand*{\oldepsilon}{\epsilon}%
\renewcommand*{\epsilon}{\varepsilon}%

% Number Sets
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\1}{\mathbf{1}}

% Statistics
\newcommand{\Var}{\mathbb{V}}
\newcommand{\Exp}{\mathbb{E}}

% Matrices
\newcommand{\X}{\mathbf{X}}
\newcommand{\Y}{\mathbf{Y}}

% operators
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\di}{d}

\begin{document}

\section{Ontology}
The constructs used are
\begin{itemize}
 \item performance
 \item performance accessment
 \item skill
 \item confidence
\end{itemize}

\section{Mathematical Formulation}

\section{(Statistical) Models}

\section{Open Questions / Ambiguities}

\end{document}