### 2025-02-26 dk_example/figures/error_dk.tex

\begin{tikzpicture}
        [manifest/.style={rectangle,draw=black!80,semithick,minimum width=1cm,inner sep=4pt,text centered},
        latent/.style={circle,draw=black!80,thick,inner sep=2,text centered},
        latent_wide/.style={circle,draw=black!80,thick,inner sep=4,text centered},
        from/.style={<-,>=stealth',semithick},
        text height=1.75ex, text depth=.5ex]

\begin{scope}

% types
\node[manifest] at (0,0) (skill) {\texttt{skill}};

\node[manifest] at (6, 0) (overconfidence) {\texttt{overconfidence}};

\node[manifest] at (0, -2) (performance) {\texttt{performance}};

\node[manifest] at (6,-2) (estimate) {\texttt{overestimation}};

\node[manifest] at (0, -6) (skillerror) {\texttt{skill + error}};

\node[manifest] at (6, -4) (pminusp) {\texttt{expected performance - performance}};

\node[manifest] at (6, -6) (ocerror) {\texttt{overconfidence - error}};

%arrows
\draw[-,>=stealth',semithick] (skill) -- node [fill=white] {$-$} (overconfidence);
\draw[-,>=stealth',semithick] (performance) -- node [fill=white] {$-$} (estimate);
\draw[-,>=stealth',semithick] (skill) -- node [fill=white] {$\simeq$} (performance);
\draw[-,>=stealth',semithick] (overconfidence) -- node [fill=white] {$\simeq$} (estimate);
\draw[-,>=stealth',semithick] (skillerror) -- node [fill=white] {$-$} (ocerror);
\draw[-,>=stealth',semithick] (performance) -- node [fill=white] {$\coloneq$} (skillerror);
\draw[-,>=stealth',semithick] (estimate) -- node [fill=white] {$\coloneq$} (pminusp);
\draw[-,>=stealth',semithick] (pminusp) -- node [fill=white] {$=$ (derived below)} (ocerror);

\end{scope}

\end{tikzpicture}